---
title: "Model Data"
date: "`r Sys.Date()`"
editor_options:
  chunk_output_type: console
output: pdf_document
---

# Load libraries

```{r}
rm(list=ls())
options (scipen = 999)
```

# Load data

`load_loans.R` was copied from https://github.com/KivaAnalysis/data.
Use a direct link when that repo goes public.

```{r}
source('./load_loans.R')
data <- load_loans('../data/loans')
rm(load_loans)
```

# Overall Ranges

Calculate any range we want over the full dataset.

```{r}
funded_percent <- 100 * xtabs(~data$status)/nrow(data)
funded_percent <- unname(funded_percent['funded'])
```

# Filter Columns

Keep only the columns of concern.

```{r}
my_vars <- c('use', 'description', 'status', 'fundraisingDate', 'raisedDate')
data <- data[,my_vars]
rm(my_vars)
```

# Filter Rows

We are not concerned with whether or not an individual loan gets funded.
This is because the overall funding rate for Kiva is so high (~`r round(funded_percent, 1)`%).

```{r}
data <- data[data$status == 'funded',]
```

Instead it is more appropriate for the target to be the speed at which funding occurs.

# QA

Filter out observations where we can not calculate our target.

* Missing dates
* Dates out of order

```{r}
data <- data[!is.na(data$fundraisingDate),]
data <- data[!is.na(data$raisedDate),]
data <- data[data$fundraisingDate <= data$raisedDate,]
```

# Calculated Columns

Calculate the _median_ funding time and use that as the split target

```{r}
data$tte <- as.numeric(as.duration(data$raisedDate - data$fundraisingDate))
funded_tte <- quantile(data$tte, probs = c(.50, .95))
data$target <- data$tte > unname(funded_tte['50%'])
```

# Summary statistics

Collect some summary stats for use throughout the paper.

```{r}
funded_min_date <- date(min(data$fundraisingDate))
funded_max_date <- date(max(data$fundraisingDate))
funded_total_loans <- nrow(data)
```

# Save

Save the working measures + target as well as the summary stats

```{r}
data$status <- NULL
data$fundraisingDate <- NULL
data$raisedDate <- NULL
data$tte <- NULL

write.csv(data, file = '../data/measures.csv', row.names = F)

save(
  funded_percent,
  funded_min_date,
  funded_max_date,
  funded_total_loans,
  funded_tte,
  file = '../results/model_data.RData')
```
